---
title: "Sampling Bias Tests"
author: "Sheryl Murdock"
date: "25 November 2020"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: cerulean
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)
opts_knit$set(root.dir = "~/Desktop/3M_for_github")
```

This page details statistical tests for compositional biases related to methods of microbial assemblage sampling. We compare between methods of cell removal from tubeworm-associated assemblages (direct removal of biofilms or agitation in a bucket of sterile artificial seawater) and microbial size fraction (micro=20-64µm, pico/nano=0.2-20µm).

# Data Import

Import data from all 40 individual DNA extracts and reduce down to the 30 from grab samples (two size fractions from each of 13 samples plus two methodological replicates from samples EMw1 and EMw6). Singleton OTUs have been removed. 
```{r}
# Bacteria
Bact40<-read.csv("Bact40nonsingleton.csv", header=TRUE, row.names = 1)
Bact30<-Bact40[-c(3,10,15,20,25,28,33,38,39,40),]

# Eukarya
Euk40<-read.csv("Euk40nonsingleton.csv", header=TRUE, row.names = 1)
Euk30<-Euk40[-c(3,10,15,20,25,28,33,38,39,40),]

# Archaea (two extracts of ECw11 removed- too few reads)
Arch38<-read.csv("Arch38nonsingleton.csv", header=TRUE, row.names = 1)
Arch28<-Arch38[-c(3,8,13,18,23,26,31,36,37,38),]

# qPCR-balanced microbial assemblage (size fractions combined, EMw1 and EMw6 replicates included)
# Remove fluid samples.
Micro25<-read.csv("QPCR_balanced_microbes_25.csv", header=TRUE, row.names = 1)
Micro25<-as.data.frame(t(Micro25))
balMicGrabs<-Micro25[-c(1:10),]

# metadata-size fractions separate
grabs30meta<-read.csv("grabs30metadata.csv", header=TRUE, row.names = 1)
# metadata-size fractions combined
grabs15meta<-read.csv("grabs15metadata.csv", header=TRUE, row.names = 1)
```
<br>

# Size fraction bias tests 
## Bacteria
Test for difference in diversity using Wilcoxon Rank Sum test.
```{r message=FALSE, warning=FALSE}
library(vegan)
bact30Div<-diversity(Bact30, index="invsimpson", MARGIN = 1)
bact30Div

grabs30meta<-cbind(grabs30meta,bact30Div)

mDivB<-grabs30meta$bact30Div[grabs30meta$Size_class=="micro"]
pnDivB<-grabs30meta$bact30Div[grabs30meta$Size_class=="pico_nano"]

wilcox.test(mDivB,pnDivB)
```
**RESULT: Bacterial diversity values are not significantly different between size fractions**
<br>

Test for difference in composition using Analysis of Similarities (ANOSIM).
```{r warning=FALSE, message=FALSE}
library(zCompositions)
library(vegan)
library(compositions)

# remove zero sum OTUs (i.e. OTUs that were only in the fluids)
Bact30trim<-Bact30[,which(colSums(Bact30) >0)]

# 'zCompositions' for imputation of zeros. 
Bact30_cmultRepl<-cmultRepl(Bact30trim, method="CZM", output="p-counts")

# Centered log-ratio transformation
clr_Bact30<-apply(Bact30_cmultRepl, 2, function(x){log(x)-mean(log(x))})

# Calculate Aitchison distance matrix
clr_Bact30.dist<-dist(clr_Bact30)

codaBact30ANOSIMsize<-with(grabs30meta, anosim(clr_Bact30.dist, Size_class))
summary(codaBact30ANOSIMsize)
```
**RESULT: Bacterial composition does NOT differ significantly between size fractions**
<br>

## Eukarya
Test for difference in diversity using Wilcoxon Rank Sum test.
```{r warning=FALSE, message=FALSE}
euk30Div<-diversity(Euk30, index="invsimpson", MARGIN = 1)
euk30Div

grabs30meta<-cbind(grabs30meta,euk30Div)

mDivE<-grabs30meta$euk30Div[grabs30meta$Size_class=="micro"]
pnDivE<-grabs30meta$euk30Div[grabs30meta$Size_class=="pico_nano"]

wilcox.test(mDivE,pnDivE)
```
**RESULT: Microeukaryote diversity values are not significantly different between size fractions**
<br>

Test for difference in composition using ANOSIM.
```{r warning=FALSE, message=FALSE}
# remove zero sum OTUs (i.e. OTUs that were only in the fluids)
Euk30trim<-Euk30[,which(colSums(Euk30) >0)]

# 'zCompositions' for imputation of zeros. 
Euk30_cmultRepl<-cmultRepl(Euk30trim, method="CZM", output="p-counts")

# Centered log-ratio transformation
clr_Euk30<-apply(Euk30_cmultRepl, 2, function(x){log(x)-mean(log(x))})

# Calculate Aitchison distance matrix
clr_Euk30.dist<-dist(clr_Euk30)

codaEuk30ANOSIMsize<-with(grabs30meta, anosim(clr_Euk30.dist, Size_class))
summary(codaEuk30ANOSIMsize)
```
**RESULT: Microeukaryote composition does NOT differ significantly between size fractions**
<br>

## Archaea
Test for difference in diversity using Wilcoxon Rank Sum test.
```{r warning=FALSE, message=FALSE}
arch28Div<-diversity(Arch28, index="invsimpson", MARGIN = 1)
arch28Div

grabs28meta<-grabs30meta[-(7:8),]
grabs28meta<-cbind(grabs28meta,arch28Div)

mDivA<-grabs28meta$euk30Div[grabs28meta$Size_class=="micro"]
pnDivA<-grabs28meta$euk30Div[grabs28meta$Size_class=="pico_nano"]

wilcox.test(mDivA,pnDivA)
```
**RESULT: Archaeal diversity values are not significantly different between size fractions**
<br>

Test for difference in composition using ANOSIM.
```{r warning=FALSE, message=FALSE}
# remove zero sum OTUs (i.e. OTUs that were only in the fluids)
Arch28trim<-Arch28[,which(colSums(Arch28) >0)]

# 'zCompositions' for imputation of zeros. 
Arch28_cmultRepl<-cmultRepl(Arch28trim, method="CZM", output="p-counts")

# Centered log-ratio transformation
clr_Arch28<-apply(Arch28_cmultRepl, 2, function(x){log(x)-mean(log(x))})

# Calculate Aitchison distance matrix
clr_Arch28.dist<-dist(clr_Arch28)

codaArch28ANOSIMsize<-with(grabs28meta, anosim(clr_Arch28.dist, Size_class))
summary(codaArch28ANOSIMsize)
```
**RESULT: Microeukaryote composition does NOT differ significantly (at p<0.01) between size fractions**

# Methodological replicate bias tests
## Composition
### Two replicate samples
Using the QPCR-balanced microbial assemblages
```{r warning=FALSE, message=FALSE}
# Test only the two paired samples, EMw1 and EMw6
reps<-balMicGrabs[c(2,3,5,6),]
reps<-reps[,which(colSums(reps) >0)]
repsMeta<-(grabs15meta[c(2,3,5,6),])

# Deal with zeros
reps_cmultRepl<-cmultRepl(reps, method="CZM", output="p-counts")

# Centered log-ratios
clr_reps <- apply(reps_cmultRepl, 2, function(x){log(x)-mean(log(x))})

# Distance matrix
clr_reps.dist<-dist(clr_reps)

reps_ANOSIM<-with(repsMeta, anosim(clr_reps.dist, Sample_type))
summary(reps_ANOSIM)
```
**RESULT: No signficant difference in microbial assemblage composition between cell harvesting methods.**
<br>

### All grab samples (QPCR-balanced)
Double check by comparing all grab samples to verify that no distinctions occur between the two methods.
```{r warning=FALSE, message=FALSE}
# Deal with zeros
balMicGrabs_cmultRepl<-cmultRepl(balMicGrabs, method="CZM", output="p-counts")

# Centered log-ratios
clr_balMicGrabs <- apply(balMicGrabs_cmultRepl, 2, function(x){log(x)-mean(log(x))})

# Distance matrix
clr_balMicGrabs.dist<-dist(clr_balMicGrabs)

balMicGrabs_ANOSIM<-with(grabs15meta, anosim(clr_balMicGrabs.dist, Sample_type))
summary(balMicGrabs_ANOSIM)
```
**RESULT: Full set of grab samples confirms, no significant compositional difference between cell harvesting methods.**

### All grab samples (Bacteria)
```{r warning=FALSE, message=FALSE}
codaBact30ANOSIMmethod<-with(grabs30meta, anosim(clr_Bact30.dist, Sample_type))
summary(codaBact30ANOSIMmethod)
```
**RESULT: Weak difference in bacterial compositions (p<0.05).**

### All grab samples (Eukarya)
```{r warning=FALSE, message=FALSE}
codaEuk30ANOSIMmethod<-with(grabs30meta, anosim(clr_Euk30.dist, Sample_type))
summary(codaEuk30ANOSIMmethod)
```
**RESULT: Very weak difference in microeukaryote compositions (p<0.05).**

### All grab samples (Archaea)
```{r warning=FALSE, message=FALSE}
codaArch28ANOSIMmethod<-with(grabs28meta, anosim(clr_Arch28.dist, Sample_type))
summary(codaArch28ANOSIMmethod)
```
**RESULT: Weak difference in archaeal compositions (p<0.05).**
<br>

## Diversity
### Bacteria
Test for difference in diversity using Wilcoxon Rank Sum test
```{r}
bkDivB<-grabs30meta$bact30Div[grabs30meta$Sample_type=="bk"]
dDivB<-grabs30meta$bact30Div[grabs30meta$Sample_type=="d"]

wilcox.test(dDivB,bkDivB)
```
**RESULT: Bacterial diversity values are NOT significantly different between the two harvesting methods**

### Euakrya
```{r}
bkDivE<-grabs30meta$euk30Div[grabs30meta$Sample_type=="bk"]
dDivE<-grabs30meta$euk30Div[grabs30meta$Sample_type=="d"]

wilcox.test(dDivE,bkDivE)
```
**RESULT: Microeukaryote diversity values ARE significantly different between the two harvesting methods**

### Archaea
```{r}
bkDivA<-grabs28meta$arch28Div[grabs28meta$Sample_type=="bk"]
dDivA<-grabs28meta$arch28Div[grabs28meta$Sample_type=="d"]

wilcox.test(dDivA,bkDivA)
```
**RESULT: Archaeal diversity values are NOT significantly different between the two harvesting methods**

# Session Info
```{r}
sessionInfo()
```

